Magnetized accretion disks around Kerr black holes with scalar hair - I. Constant angular momentum disks

### 2018-11-17 paper_prd.tex

The models of KBHsSH we use in this study are built following the procedure described in~\cite{Herdeiro:2015b}. The underlying theoretical framework is the Einstein-Klein-Gordon (EKG) field theory, describing a massive complex scalar field $\Psi$ minimally coupled to Einstein gravity. KBHsSH solutions are obtained by using the following ansatz for the metric and the scalar field~\cite{Herdeiro:2014a}
\begin{eqnarray}
%\begin{split}
\mathrm{d}s^2 &=& e^{2F_1}\left(\frac{\mathrm{d}r^2}{N} + r^2\mathrm{d}\theta^2\right) +  e^{2F_2}r^2\sin^2 \theta(\mathrm{d}\phi-W\mathrm{d}t)^2 
\nonumber \\ 
&-&  e^{2F_0}N\mathrm{d}t^2\,,
\label{metric}
\end{eqnarray}
\begin{eqnarray}
\Psi = \phi(r, \theta) e^{\mathrm{i}(m\varphi - \omega t)} \,,
%\end{split}
\end{eqnarray}
with $N = 1 - r_{\rm H}/r$, where $r_{\rm H}$ is the radius of the event horizon of the BH, and $W$, $F_1$, $F_2$, $F_0$ are functions of $r$ and $\theta$. Moreover, $\omega$ is the scalar field frequency and $m$ is the azimuthal harmonic index.
We note that the radial coordinate $r$ is related to the Boyer-Lindquist radial coordinate $r_{\rm BL}$ by $r=r_{\rm BL}-a^2/r_{\rm H,BL}$, where $a$ stands for the spin of the BH and $r_{\rm H,BL}$ is the location of the horizon in Boyer-Lindquist coordinates.

@@

To account for the magnetic field in the disks we use the procedure described by~\cite{Komissarov:2006,Montero:2007}. First, we write the equations of ideal general relativistic MHD as the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (\rho h + b^2)u^{\mu}u^{\nu} + (p + p_{\mathrm{m}})g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetized perfect fluid, $h$, $\rho$, $p$, and $p_{\mathrm{m}}$ being the fluid specific enthalpy, density, fluid pressure, and magnetic pressure, respectively, the latter defined as $p_{\mathrm{m}} = b^2/2$. The ratio of fluid pressure to magnetic pressure defines the magnetization parameter $\beta_{\mathrm{m}} = p/p_{\mathrm{m}}$.
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$ (see~\cite{Anton:2006} for further details). Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting the divergence of Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(\rho h + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. This equation can be rewritten in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, 
\begin{equation}\label{eq:diff_ver}
\partial_i(|\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{\rho h} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}\rho h} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.

@@

In figures~\ref{models_peri_I} and~\ref{models_peri_II} we show the same morphological distribution of Figs.~\ref{models_I} and~\ref{models_II} but using, instead, a perimetral radial coordinate $R$, related to the radial coordinate $r$ according to $R = e^{F_2} r$. This perimetral coordinate represents the proper length along the azimuthal direction, which constitutes a geometrically meaningful direction since it runs along the orbits of the azimuthal Killing vector field. Therefore, the proper size of a full $\phi$ orbit is given by $2\pi R$, i.e.~$R$ is the perimetral radius. The most salient feature of the morphologies shown in Figs.~\ref{models_peri_I} and~\ref{models_peri_II}, when comparing to those displayed in Figs.~\ref{models_I} and~\ref{models_II}, is the deformation of the disks in their innermost regions. In general, the deformations become larger the higher the horizon sphericity $\mathfrak{s}$ and the closer the disk is to the horizon. Model III is the one showing the largest deformation, as $(R_{\rm in}-R_{\rm H}/R_{\rm H}$ attains the smallest value for this model. \tf{Check.} \sg{Done} It is also worth noticing that the shape of the BH also changes when using the perimetral coordinate. While in the $r$ coordinate the horizon is spherical (cf.~Figs.~\ref{models_I} and~\ref{models_II}) in the perimetral coordinate $R$ is not always so. Moreover, the larger the value of $v_{\rm H}$, the more elliptic the horizon becomes, which in our sample corresponds to model III (cf.~Table~\ref{models_list}, $\mathfrak{s}=1.489$). Also, an interesting property of the perimeteral coordinates is that. for the Kerr metric, $R_{\mathrm{H}} = 2M$ irrespective of the value of the angular momentum. But for the KBHsSH cases, $2M_{\mathrm{H}} \leq R_{\mathrm{H}} \geq 2M_{\mathrm{ADM}}$, and the quotient $R_{\mathrm{H}}/ 2M_{\mathrm{H}}$ increases as more mass and angular momentum is stored in the scalar field. \sg{It is possible that the angular momentum does not change anything, but we cannot know, as we do not have two different models with the same value of the BH mass.}

\sg{Discuss. I find that the 2D morphology in perimeteral coordinates might not be very useful, but I found an interesting property of these coordinates: For the Kerr metric, $R_{\mathrm{H}} = 2M$ irrespective of the value of the angular momentum (For the KBHsSH cases, $2M_{\mathrm{H}} \leq R_{\mathrm{H}} \geq 2M_{\mathrm{ADM}}$. I don't know if there are any kind of dependence on the angular momentum, as we don't two cases with the same value of the mass). I suppose this could help us to compare different models. In fact, the differences in the morphology close to the equatorial plane in the near horizon region are actually related to the value of the angular momentum (we can see a relationship between the morphological differences and the equivalent spin parameter $a_{\mathrm{eq}}$).} \tf{Could you please include some of this information in here?} \sg{Done}

>>>>

The models of KBHsSH we use in this study are built following the procedure described in~\cite{Herdeiro:2015b}. The underlying theoretical framework is the Einstein-Klein-Gordon (EKG) field theory, describing a massive complex scalar field $\Psi$ minimally coupled to Einstein gravity. KBHsSH solutions are obtained by using the following ansatz for the metric and the scalar field~\cite{Herdeiro:2014a}
\begin{eqnarray}
%\begin{split}
\mathrm{d}s^2 &=& e^{2F_1}\left(\frac{\mathrm{d}r^2}{N} + r^2\mathrm{d}\theta^2\right) +  e^{2F_2}r^2\sin^2 \theta(\mathrm{d}\phi-W\mathrm{d}t)^2 
\nonumber \\ 
&-&  e^{2F_0}N\mathrm{d}t^2\,,
\label{metric}
\end{eqnarray}
\begin{eqnarray}
\Psi = \phi(r, \theta) e^{\mathrm{i}(m\varphi - \omega t)} \,,
%\end{split}
\end{eqnarray}
with $N = 1 - r_{\rm H}/r$, where $r_{\rm H}$ is the radius of the event horizon of the BH, and $W$, $F_1$, $F_2$, $F_0$ are functions of $r$ and $\theta$. Moreover, $\omega$ is the scalar field frequency and $m$ is the azimuthal harmonic index.
We note that the radial coordinate $r$ is related to the Boyer-Lindquist radial coordinate $r_{\rm BL}$ by $r=r_{\rm BL}-a^2/r_{\rm H,BL}$, where $a = J/M$ stands for the spin of the BH and $r_{\rm H,BL}$ is the location of the horizon in Boyer-Lindquist coordinates.

@@

To account for the magnetic field in the disks we use the procedure described by~\cite{Komissarov:2006,Montero:2007}. First, we write the equations of ideal general relativistic MHD as the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (\rho h + b^2)u^{\mu}u^{\nu} + (p + p_{\mathrm{m}})g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetized perfect fluid, $h$, $\rho$, $p$, and $p_{\mathrm{m}}$ being the fluid specific enthalpy, density, fluid pressure, and magnetic pressure, respectively, the latter defined as $p_{\mathrm{m}} = b^2/2$. The ratio of fluid pressure to magnetic pressure defines the magnetization parameter $\beta_{\mathrm{m}} = p/p_{\mathrm{m}}$.
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$ (see~\cite{Anton:2006} for further details). Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting the divergence of Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(\rho h + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. This equation can be rewritten in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, 
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln |u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{\rho h} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}\rho h} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.

@@

In figures~\ref{models_peri_I} and~\ref{models_peri_II} we show the same morphological distribution of Figs.~\ref{models_I} and~\ref{models_II} but using, instead, a perimetral radial coordinate $R$, related to the radial coordinate $r$ according to $R = e^{F_2} r$. This perimetral coordinate represents the proper length along the azimuthal direction, which constitutes a geometrically meaningful direction since it runs along the orbits of the azimuthal Killing vector field. Therefore, the proper size of a full $\phi$ orbit is given by $2\pi R$, i.e.~$R$ is the perimetral radius. The most salient feature of the morphologies shown in Figs.~\ref{models_peri_I} and~\ref{models_peri_II}, when comparing to those displayed in Figs.~\ref{models_I} and~\ref{models_II}, is the deformation of the disks in their innermost regions. In general, the deformations become larger the higher the horizon sphericity $\mathfrak{s}$ and the closer the disk is to the horizon. Model III is the one showing the largest deformation, as $R_{\rm in}-R_{\rm H}/R_{\rm H}$ attains the smallest value for this model. \tf{Check.} \sg{Done} It is also worth noticing that the shape of the BH also changes when using the perimetral coordinate. While in the $r$ coordinate the horizon is spherical (cf.~Figs.~\ref{models_I} and~\ref{models_II}) in the perimetral coordinate $R$ is not always so. Moreover, the larger the value of $v_{\rm H}$, the more elliptic the horizon becomes, which in our sample corresponds to model III (cf.~Table~\ref{models_list}, $\mathfrak{s}=1.489$). Also, an interesting property of the perimeteral coordinates is that, for the Kerr metric, $R_{\mathrm{H}} = 2M$ irrespective of the value of the angular momentum. But for the KBHsSH cases, $2M_{\mathrm{H}} < R_{\mathrm{H}} < 2M_{\mathrm{ADM}}$, and the quotient $R_{\mathrm{H}}/ 2M_{\mathrm{H}}$ increases as more mass and angular momentum is stored in the scalar field. \sg{It is possible that the angular momentum does not change anything, but we cannot know, as we do not have two different models with the same value of the BH mass.}

\sg{Discuss. I find that the 2D morphology in perimeteral coordinates might not be very useful, but I found an interesting property of these coordinates: For the Kerr metric, $R_{\mathrm{H}} = 2M$ irrespective of the value of the angular momentum (For the KBHsSH cases, $2M_{\mathrm{H}} < R_{\mathrm{H}} < 2M_{\mathrm{ADM}}$. I don't know if there are any kind of dependence on the angular momentum, as we don't two cases with the same value of the mass). I suppose this could help us to compare different models. In fact, the differences in the morphology close to the equatorial plane in the near horizon region are actually related to the value of the angular momentum (we can see a relationship between the morphological differences and the equivalent spin parameter $a_{\mathrm{eq}}$).} \tf{Could you please include some of this information in here?} \sg{Done}
